---
title: "linear model from Web"
author: "Logan DeHay"
date: "Fall 2021"
output:
  pdf_document:
    toc: yes
    toc_depth: '5'
  html_document:
    toc: yes
    toc_depth: 5
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Read about the data from the website where it is staged.  Then form three regression models; Model1 is SIMS~ARM, Model2 is SIMS~GRIP, and Model3 is SIMS~ARM+GRIP. For each model find a 95% prediction interval of SIMS given a value of 94  for GRIP and 88 for ARM. Compare Model1 with Model3 using anova. Write it up in a markdown document, push the project up to your github account and submit it back to canvas as link. 

 


```{r include=FALSE}
require(tidyverse)
require(tigerstats)
require(rgl)
require(knitr)

```

```{r}
data <- read.csv(file="https://onlinestatbook.com/case_studies_rvls/physical_strength/data.txt",sep="",header=TRUE)  

```

## Model 1 SIM~ARMS

### scatterplot
  
```{r}
plot(SIMS~ARM,data=data)
```
  
The ARM v. SIMS scatter plot illustrates a positive correlation between the two variables. The scatter plot is fairly tight with a relatively consistent band width of about 4. It can be estimated that a standard deviation of the data is equal to approximately 1.
  
### Numerical results
  
```{r}
cor(SIMS~ARM,data=data)
```
  
### Inferential  (Build model.1)
  
```{r}
model.1 <- lm(SIMS~ARM,data=data)
summary.lm(model.1)
```
  
SIMS= 0.054563(ARM) - 4.095160
Our adjusted R-squared value is 0.467, which is indicating that ARM explains less than half the variation of the SIMS data. The addition of new regressors induces a decrease in the value of adjusted R-squared. It reduces the error of the mean model of the equation by 46.7%. In addition the calculated residual standard error of the data is 1.226, which is a measure of the deviation from our regression line. According to this data our equation of the line is SIMS=0.054563*(ARM)-4.095160. 
  
#### Predict at target point
```{r}
SIMS=0.054563*(88)-4.095160
```

  
```{r}
mydata= data.frame(GRIP=94,ARM=88)
```

```{r}
predict(model.1,mydata,interval="prediction")
```
  
The prediction interval is drastically larger than the confidence interval ranging from -1.72 to 3.13. This is equal to about 4 residual standard errors. When ARM is equal to 88, SIMS is predicted to be 0.7063836. 
  
#### Scatterplot with model fit
  
```{r}
basicNN <- ggplot(data,aes(y=SIMS,x=ARM))
```
  
```{r}
basicNN + geom_point() + geom_smooth(method=lm)
``` 
  
There is a normal curve above every point on the x-axis and the peak of the curve is center on the blue line. Each side of the line displays errors.
  
## Model 2 SIM~GRIP

### Now add in scatterplot
```{r}
plot(SIMS~GRIP,data=data)
```
  
There is a general trend upwards and to the right which indicates a positive linear correlation between GRIP and SIMS.
  
### Numerical results 
```{r}
cor(SIMS~GRIP,data=data)
```
  
The correlation of SIMS and GRIP is 0.6398 which is lesser than the correlation between SIMS and ARM of 0.6830. Despite this, both are strong correlations.
  
### Inferential  (Build model.2)
  
```{r}
model.2 <- lm(SIMS~GRIP,data=data)
summary.lm(model.2)
```
  
The equation of the line is SIMS= 0.045463*(GRIP)-4.809675
  
#### predict model.2 at target point

```{r}
predict(model.2,mydata,interval="prediction")
```
  
This prediction interval is wider than the same interval for SIMS and ARM. When grip is 94, then SIMS is predicted to be -0.5361543.
  
#### now add the model fit to our plot for model.2
  
```{r}
basicNT <- ggplot(data,aes(y=SIMS,x=GRIP))
```
  
```{r}
basicNT + geom_point() + geom_smooth(method=lm)
``` 

## Model 3 SIM~ARM+GRIP

```{r}
model.3 <- lm(SIMS~ARM + GRIP,data=data)
summary.lm(model.3)
``` 
  
SIMS= (0.037311* ARM)+(0.024470* GRIP)- 5.433871
  
### Numerical results (cor)
  
```{r}
cor(SIMS~ARM+GRIP,data=data)
```
  
This correlation is greater than either ARM or GRIP. The correlation is 73.27%.
  
### Inferential  (Build 2-dimentional model.3)
```{r}
model.3 <- lm(SIMS~ARM+GRIP,data=data)
summary.lm(model.3)
```
  
SIMS= (0.037311* ARM)+ (0.024470* GRIP)- 5.433871
  
#### predict model.3 at target point
  
```{r}
predict(model.3,mydata,interval="prediction")
```
  
When ARM is equal to 88 and GRIP equal to 94, the value of SIMS is estimated to be 0.1496476. Interval ranges from -2.13 to 2.43.
  
## Comparing nested models ANOVA Test

### Model.1 vs Model.3
```{r}
anova(model.1,model.3)
```
  
The test displays a difference between the accuracy of the two models. The p-value is 4.994e-06, and such a small value strongly suggests the likelihood that the addition of GRIP had a significant impact on the improvement of the model. The residual sum of errors decreased from model1 to model2. The addition of GRIP to model.1 decreased the errors from 217.88 to 188.43. The model with two regressors in it is superior in accuracy comparatively to model.1 which only has one.
  
### Model.2 vs Model.3
```{r}
anova(model.2,model.3)
```
  
Model.2 has an even greater number of errors than model.1 and as a result, the addition of ARM to model.2 has a greater impact on the accuracy of the model. The p-value is extremely small being 1.495 nano-units. This indicates an even greater probability of the addition of ARM to Model.2 has a significant impact on the accuracy of the model.
  
## Informally compare Model.1 with model.2
```{r}
anova(model.1,model.2)
```
  
Model.1 is more accurate than Model.2 which is evident from the smaller residual sum of squares. However, a p-value is not available due to their status as nested models.
  