---
title: "linear model from Web"
author: "Logan DeHay"
date: "Fall 2021"
output:
   html_document:
         toc: true
         toc_depth: 5
        
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Read about the data from the website where it is staged.  Then form three regression models; Model1 is SIMS~ARM, Model2 is SIMS~GRIP, and Model3 is SIMS~ARM+GRIP. For each model find a 95% prediction interval of SIMS given a value of 94  for GRIP and 88 for ARM. Compare Model1 with Model3 using anova. Write it up in a markdown document, push the project up to your github account and submit it back to canvas as link. 

 


```{r include=FALSE}
require(tidyverse)
require(tigerstats)
require(rgl)
require(knitr)

```

```{r}
data <- read.csv(file="https://onlinestatbook.com/case_studies_rvls/physical_strength/data.txt",sep="",header=TRUE)  

```

## Model 1 SIM~ARMS

### scatterplot
  
```{r}
plot(SIMS~ARM,data=data)
```
  
The ARM v. SIMS scatter plot illustrates a positive correlation between the two variables. The scatter plot is fairly tight with a relatively consistent band width of about 4. It can be estimated that a standard deviation of the data is equal to approximately 1.
  
### Numerical results
  
```{r}
cor(SIMS~ARM,data=data)
```
  
### Inferential  (Build model.1)
  
```{r}
model.1 <- lm(SIMS~ARM,data=data)
summary.lm(model.1)
```
  
SIMS= 0.054563(ARM) - 4.095160
Our adjusted R-squared value is 0.467, which is indicating that ARM explains less than half the variation of the SIMS data. It reduces the error of the mean model of the equation by 46.7%. In addition the calculated residual standard error of the data is 1.226, which is a measure of the deviation from our regression line. According to this data our equation of the line is SIMS=0.054563*(ARM)-4.095160. 
  
#### Predict at target point
```{r}
SIMS=0.054563*(88)-4.095160
```

  
```{r}
mydata= data.frame(GRIP=94,ARM=88)
```

```{r}
predict(model.1,mydata,interval="confidence")
```
  
#### Scatterplot with model fit
  
```{r}
basicNN <- ggplot(data,aes(y=SIMS,x=ARM))
```
  
```{r}
basicNN + geom_point() + geom_smooth(method=lm)
``` 
  
## Model 2 SIM~GRIP

### Now add in scatterplot


### Numerical results 


### Inferential  (Build model.2)
  

#### predict model.2 at target point

#### now add the model fit to our plot for model.2
  


## Model 3 SIM~ARM+GRIP

### Numerical results (cor)

`
  
### Inferential  (Build 2-dimentional model.3)

#### predict model.3 at target point
  

## Comparing nested models ANOVA Test

### Model.1 vs Model.3
### Model.2 vs Model.3

## Informally compare Model.1 with model.2

